### 2016-06-11 conclusions.tex
Nueva organizacion de los archivos.  Partes del paper en archivos separados y todas las imagenes en subdirectorio fig

\section{Conclusions}
\label{sec:concl}

\subsection{Future work}