### 2017-01-16 conclusions.tex
2017-01-13-prueba

\section{Conclusions}
\label{sec:concl}

\subsection{Future work}
temporal