### 2017-05-02 conclusions.tex
1-5 6.59

\section{Conclusions}
\label{sec:concl}
Este estudio presentó la comparación de varias técnicas de machine learning en la predicción de the la tasa de desarrollo de la Sigatoka negra. Al respecto se presentan las conclusionez al respecto:
\begin{itemize}
\item 
\item 
\item 
\item

\end{itemize}

A comparison of machine learning techniques
The main findings were: 1) The highest R2 was 60\% 2) The highest R2 were reached with linear models like support vector regression with linear kernel, 3) As little as three meteorological variables can be used because of the correlations detected among variables.

\subsection{Future work}
Currently this research is focused on the extension of these findings to other farms and the extension of the model to be applied in an early warning system, which does not require the regression task.